---
title: "NYPD SHOOTING ANALYSIS"
author: "TT"
# date: "2024-05-02"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r downloaddata, include=TRUE}
library(tidyverse)
nypd <- read_csv("https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv")
View(nypd)
```

### Tidy the Data
```{r cleandata, include=TRUE}
columns_to_keep <- c("INCIDENT_KEY", "BORO", "PERP_SEX", "PERP_AGE_GROUP")
filtered_data <- nypd[, columns_to_keep, drop = FALSE] %>%
  filter(!rowSums(is.na(select(., "PERP_SEX"))) & 
  !rowSums(select(., "PERP_SEX") == "(null)"))


View(filtered_data)
```

**Models**

My model will be to see if there are any correlation between age_group and gender with the rate of gunshots.

**Identify any Bias**

My bias: 
- I think there will be more Male compared to Females.
- I think there will be more younger (under 25) people compared to over 25.

## visualization 1

```{r showvisual1, include=TRUE}
ggplot(filtered_data, aes(x = PERP_SEX)) +
  geom_bar() +
  labs(x = "Gender", y = "Count", title = "Count by Gender")
```

## visualization 2

```{r showvisual2, include=TRUE}
ggplot(filtered_data, aes(x = PERP_AGE_GROUP)) +
  geom_bar() +
  labs(x = "Gender", y = "Count", title = "Count by AGE GROUP")
```

## clean up the data form AGE GROUP
remove 1020, 1028, 224, 940, UNKNOWN, NA
```{r cleandataagain, include=TRUE}
values_to_remove <- c("1020", "1028", "224", "940", "UNKNOWN", "NA")

filtered_age <- filtered_data %>%
  filter(!PERP_AGE_GROUP %in% values_to_remove)
ggplot(filtered_age, aes(x = PERP_AGE_GROUP)) +
  geom_bar() +
  labs(x = "Gender", y = "Count", title = "Count by AGE GROUP")
```

## Summary
- I think that there is a higher number of Males compared to Females in the result.
- I think my 2nd visualization does not make a clear distinction whether there are more Perps that are under the age of 24 compared to above it.


```{r show_session}
"ran sessioninfo in console:
 sessionInfo()
R version 4.4.0 (2024-04-24 ucrt)
Platform: x86_64-w64-mingw32/x64
Running under: Windows 11 x64 (build 22631)

Matrix products: default


locale:
[1] LC_COLLATE=English_United States.utf8  LC_CTYPE=English_United States.utf8   
[3] LC_MONETARY=English_United States.utf8 LC_NUMERIC=C                          
[5] LC_TIME=English_United States.utf8    

time zone: Asia/Hong_Kong
tzcode source: internal

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

loaded via a namespace (and not attached):
 [1] compiler_4.4.0    fastmap_1.1.1     cli_3.6.2         htmltools_0.5.8.1 tools_4.4.0      
 [6] yaml_2.3.8        tinytex_0.50      rmarkdown_2.26    knitr_1.46        digest_0.6.35    
[11] xfun_0.43         rlang_1.1.3       evaluate_0.23" 
```
